Changeset Analysis
========================================================

Part 0: Define custom functions
```{r custom_functions, echo=FALSE, results=FALSE}
#Allow for normalization of date
adjustDate <- function(dataframe, date){
  dataframe$DaysSinceEvent = as.numeric(as.Date(dataframe$date) - as.Date(date))
  return(dataframe)
}

#This function returns 0 if x is less than zero instead of calculating the log
custom_log <- function(x) ifelse(x <= 0, 0, base::log(x, base=10)) #Redefine log to return 0 when undefined...
```

Part 0.1: Load Libraries
```{r load_libraries, echo=FALSE, results='show'}
library(ggplot2)
library(ineq)
library(plotrix)
```

Part 1: Load the Data
-------------------------------------------------------
```{r import_from_mongo, echo=FALSE, results='show', message=FALSE}
#Load Data from Mongo?  No.
#source("/Users/jenningsanderson/Dropbox/OSM/osm-history/rcode/changesets_analysis.r")

#Philippines
#phil_cursor = custom_timeboxed_cursor(ISOdate(2013,11,08), ISOdate(2013,12,08), 'philippines.changesets')
#phil_info = buildChangeset(phil_cursor)

#Haiti
#haiti_cursor = custom_timeboxed_cursor(ISOdate(2010,01,12), ISOdate(2010,02,12), 'haiti.changesets')
#haiti_info = buildChangeset(haiti_cursor)

#Combine the DFs:
#phil_info$Country  = "Philippines"
#haiti_info$Country = "Haiti"

#dat = rbind(phil_info, haiti_info)
```

1. Load the entire Data from the disaster window CSV
```{r load_entire_window, echo=FALSE, results='show', message=FALSE, warning=FALSE}

#Load the data from csv
philippines = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/phil_changesets_data.csv")
haiti       = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/haiti_changesets_data.csv")

#Combine the two now
philippines$Country  = "Philippines"
haiti$Country        = "Haiti"

#Add the date adjust
haiti = adjustDate(haiti, ISOdate(2010,1,12))
philippines = adjustDate(philippines, ISOdate(2013,11,08))

dat = rbind(philippines, haiti)
```

2. Load some small sample data, this is data from 25 random users in each data set.
```{r load_sample_data, echo=FALSE, results='show'}
phil_sample_data = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/phil_sample_data.csv")
haiti_sample_data = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/haiti_sample_data.csv")

haiti_sample_data$Country = "Haiti"
phil_sample_data$Country = "Philippines"

sample_dat = rbind(haiti_sample_data, phil_sample_data)
```

3. For more data, load a larger sample: 75 users from each set.
```{r load_sample_75, results='show', echo=FALSE}
phil_75_sample_data = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/phil_75_sample_data.csv")
haiti_75_sample_data = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/haiti_75_sample_data.csv")


haiti_75_sample_data$Country = "Haiti"
phil_75_sample_data$Country = "Philippines"

haiti_75_sample_data = adjustDate(haiti_75_sample_data, ISOdate(2010,1,12))
phil_75_sample_data = adjustDate(phil_75_sample_data, ISOdate(2013,11,08))

sample_dat75 = rbind(haiti_75_sample_data, phil_75_sample_data)
```

4. Load only changesets with less than 1000 nodes (still has the 1 node)
```{r load_1000node_sample, echo=FALSE, results='show'}
phil_lim300 = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/limited/phil_300_sample_data.csv")
haiti_lim300 = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/limited/haiti_300_sample_data.csv")

haiti_lim300$Country = "Haiti"
phil_lim300$Country = "Philippines"

haiti_lim300 = adjustDate(haiti_lim300, ISOdate(2010,1,12))
phil_lim300 = adjustDate(phil_lim300, ISOdate(2013,11,08))

sample_dat_lim = rbind(haiti_lim300, phil_lim300)

```

5. Load only changesets with less than 1000 nodes (and more than 1 node) -- entire set
```{r pruned_set, echo=FALSE, results='show'}
phil_lim = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/limited/phil_changesets_lim_data.csv")
haiti_lim = read.csv("/Users/jenningsanderson/Google Drive/OSM-Haiti-Philippines-SocialComputing/data/users_changesets/limited/haiti_changesets_lim_data.csv")

haiti_lim$Country = "Haiti"
phil_lim$Country = "Philippines"

haiti_lim = adjustDate(haiti_lim, ISOdate(2010,1,12))
phil_lim = adjustDate(phil_lim, ISOdate(2013,11,08))

dat_lim = rbind(haiti_lim, phil_lim)


```

### Summary Statistics of the data
_Currently not implemented_

Philippines
```{r summary_stats_phil, echo=FALSE, results='hide'}
summary(philippines$node_count)
summary(philippines$area)
summary(philippines$density)
```

Haiti
```{r summary_stats_haiti, echo=FALSE, results='hide'}
summary(haiti$node_count)
summary(haiti$area)
summary(haiti$density)
```

Part 2. Visual Comparison
-------------------------------------
### Comparing Node Counts
Notice that there is a large spike in the Philippines at 4 nodes per changeset.  A simple building is typically comprised of exaclty 4 nodes.  The working hypothesis here is that each of these changesets represents a building and that buildings were mapped in the Philippines because the road structure was already in place; whereas in Haiti, the road system was not yet in place.

```{r comparing_node_counts, echo=FALSE, fig.width=12, fig.height=5}
png("/Users/jenningsanderson/Dropbox/OSM_Behavioral_Analysis_CSCW/Images/ComparingNodeCounts.png",width=800, height=400)
par(mfcol=c(1,2)) 
hist(log(haiti$node_count, base=10), breaks=50, main="Nodes per changeset in Haiti", xlab="Nodes Per changeset (log base 10)")
hist(log(philippines$node_count, base=10), breaks=50, main="Nodes per changeset in Philippines",xlab="Nodes Per changeset (log base 10)")
dev.off()
```


### Lorenz curves for the amount of nodes
```{r lorenz_curves, echo=FALSE, fig.width=15}
library(ineq)
plot.new()
Lc.1 <- Lc(aggregate(haiti_lim$node_count, list(haiti_lim$user), sum)$x);
Lc.2 <- Lc(aggregate(phil_lim$node_count, list(phil_lim$user), sum)$x);
Lc.3 <- Lc(haiti_lim$user)
Lc.4 <- Lc(phil_lim$user)
plot(Lc.1, col=2, xlab="Percentage of Users", ylab="Percentage of Nodes or Changesets", main="Lorenz Curve for Changeset and Node count Ownership")
lines(Lc.2, col=4)
lines(Lc.3, lty='dotted', col=2)
lines(Lc.4, lty='dotted', col=4)

# Legends ---------------------------------------------
lines(c(.05,.15),c(.9,.9), col=2)
lines(c(.05,.15),c(.85,.85), col=4)
lines(c(.05,.15),c(.8,.8), col=2, lty='dotted')
lines(c(.05,.15),c(.75,.75), col=4, lty='dotted')
text(.15,.9,"Haiti Nodes by User", pos=4)
text(.15,.85,"Philippines Nodes by User", pos=4)
text(.15,.8,"Haiti Changesets by User", pos=4)
text(.15,.75,"Philippines Changesets by User", pos=4)
```


### Count users that edited in both sets:
```{r counting_intersect, echo=FALSE}
length(intersect(philippines$user, haiti$user))
```

### Getting an idea for WHEN the edits occur
```{r when_edits_occured, echo=FALSE, fig.width=15}
haiti_node_count = c(by(haiti$node_count, haiti$DaysSinceEvent, sum))
phil_node_count = c(by(philippines$node_count, philippines$DaysSinceEvent, sum))

height <- rbind(haiti_node_count, phil_node_count)
png("/Users/jenningsanderson/Dropbox/OSM_Behavioral_Analysis_CSCW/Images/nodes_edited_per_day.png",width=800, height=400)
barplot(height, beside = TRUE, main="Nodes Edited Each Day since Event", xlab="Days Since Event", ylab="Total number of nodes edited")
legend(x=60,y=300000,   				                 # places a legend at the appropriate place 
			 c("Philippines","Haiti"),         # puts text in the legend 
       pch=c(15,15),                   		 		 # gives the legend appropriate symbols
       col=c("grey","black"),
			 cex=1.5) 	
dev.off()
```

### Looking at Changeset Density & Area
The following results are normalized to 1, so that they can be compared side-by-side.  The breaks are .25, so the sum of the area is 4.
```{r changeset_densities_areas, echo=FALSE, message=FALSE, fig.width=12}
png("/Users/jenningsanderson/Dropbox/OSM_Behavioral_Analysis_CSCW/Images/ChangesetDensityComparison.png",width=800, height=400)
multhist(list(log(haiti_lim$density, base=10), log(phil_lim$density, base=10)), probability=TRUE, xlab="Density: Nodes/Sq. km (log base 10)", ylab="Normalized Frequency", main="Histogram of Philippines vs. Haiti Changeset Densities",breaks=seq(-2,9,by=.25))
legend(x=100,y=0.4,     			                 # places a legend at the appropriate place 
			 c("Philippines","Haiti"),         # puts text in the legend 
       pch=c(15,15),                   		 		 # gives the legend appropriate symbols
       col=c("grey","black"),
			 cex=1.5) 	
dev.off()

png("/Users/jenningsanderson/Dropbox/OSM_Behavioral_Analysis_CSCW/Images/ChangesetAreaComparison.png",width=800, height=400)
multhist(list(log(haiti_lim$area, base=10), log(phil_lim$area, base=10)), probability=TRUE, xlab="Area: Sq. km (log base 10)", ylab="Normalized Frequency", main="Histogram of Philippines vs. Haiti Changeset Areas", breaks=seq(-2.5,6,by=.25))
legend(x=70,y=0.4,     			                 # places a legend at the appropriate place 
			 c("Philippines","Haiti"),         # puts text in the legend 
       pch=c(15,15),                   		 		 # gives the legend appropriate symbols
       col=c("grey","black"),
			 cex=1.5) 	
dev.off()
#h = hist(log(haiti_lim$density))
#h$density = h$counts / sum(h$counts)*100
```

### Plot User contributions by day since the event
```{r haiti_nodes_by_user, fig.width=15}
#Load the CSV that's printed from ruby
haiti_by_day_contributions = read.csv("/Users/jenningsanderson/osm-history/ruby/csv_exports/haiti_top_20_user_node_count_by_day.csv")
ggplot(haiti_by_day_contributions, aes(x=DaysSinceEvent, y=NodeCount, color=User)) + 
  geom_line() + 
  scale_fill_brewer(palette="Paired") +
  xlab("Days Since Event") +
  ylab("Nodes Edited Each Day") +
  ggtitle("Haiti User Activity per day since Event")

```

```{r phil_nodes_by_user, fig.width=15}
#Load the CSV that's printed from ruby
phil_by_day_contributions = read.csv("/Users/jenningsanderson/osm-history/ruby/csv_exports/phil_top_20_user_node_count_by_day.csv")
ggplot(phil_by_day_contributions, aes(x=DaysSinceEvent, y=NodeCount, color=User)) + 
  geom_line() + 
  scale_fill_brewer(palette="Blues") +
  xlab("Days Since Event") +
  ylab("Nodes Edited Each Day") +
  ggtitle("Philippines User Activity per day since Event")

```
